---
title: "Most searched programming languages by tag on stackoverflow 2004-2018"
output: github_document
---

#### **Quick introduction**
____________________________________________________________________________________________________________________________________________

Anyone who has worked with programming languages knows how helpful platforms such as Stack overflow are.By using appropriate tags, we gain access to a treasure trove of knowledge which are answers of other users who struggled with similar problems.

I want to find out how often R and Python and their libraries have been searched over time. I also want to compare R and Python to other programs that enable data analysis. Also to see which languages were most searched in the period 2008 - 2018.
I hope you enjoy this simple analysis. 


#### **Data**
____________________________________________________________________________________________________________________________________________
The following analysis and visualizations are based on data downloaded from DataCamp [link](https://projects.datacamp.com/projects/435). They represent the number of searches for particular tags on the Stack Overflow platform between 2008 and 2018. 


#### **Analysis**
____________________________________________________________________________________________________________________________________________

###### Data loading

For this analysis I use the R with the packages 'tidyverse', 'lubridate', 'ggthemr'.

```{r message=FALSE, warning=FALSE, include=TRUE}
# Load libraries
library(tidyverse)
library(lubridate)
library(ggthemr)
ggthemr("dust")
```


```{r message=FALSE, warning=FALSE, include=TRUE}
# Load data
tag_by_year <- read_csv('by_tag_year.csv')

```

###### Checking data

A quick look at the data before analysis. Let's see what the table looks like and what the variable type is. 

```{r}
# check the data
print(tag_by_year)
glimpse(tag_by_year)
summary(tag_by_year)
```

The column 'year' is a 'numeric' type. Let's change it to "date" type.

```{r}
class(tag_by_year$year)
tag_by_year$year <- ymd(tag_by_year$year, truncated = 2L)
class(tag_by_year$year)

```



Let's look for missing and unique values

```{r}
# missing values 
sum(is.na(tag_by_year))

# unique values in columns tag, year_total
unique_tags <- tag_by_year %>% 
  select(tag) %>% 
  unique()

unique_tags
```

##### **Total searches over time**
____________________________________________________________________________________________________________________________________________

Before we look at the results of individual tag searches, let's analyse the number of total searches over the years. 

```{r message=FALSE, warning=FALSE, include=TRUE}
total_searches_per_year <- tag_by_year %>% 
  group_by(year, year_total) %>% 
  summarise()

total_searches_per_year
```

###### Let's visualise it

```{r}
ggplot(total_searches_per_year, aes(year, year_total)) +
  geom_line(lwd = 1, linejoin = "round") +
  ggtitle("Total searches per year") +
  xlab("Year") +
  ylab("Number of searches") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position = "none")

ggsave("myplot.png", device = "png",
       width = 10, height = 4, units = "in", dpi = 600, type = "cairo-png")
```

##### **R vs Python**
____________________________________________________________________________________________________________________________________________
Let's see how the percentage of searches for R and Python breaks down compared to all languages.

###### First, add a column with percentages

```{r}
# Add percentage column
tag_by_year_percentage <- tag_by_year %>%
  mutate(fraction = (number/year_total)*100)

# Print the new table
print(tag_by_year_percentage)
```

###### Then, filter out R and Python
```{r}
# Filter for R and python tags
r_python_over_time <- tag_by_year_percentage %>%
  filter(tag %in% c("r", "python"))

# Print the new table
print(r_python_over_time)
```

###### Visualize it

```{r}
# Create a line plot of fraction over time
ggplot(r_python_over_time, aes(year, fraction, color = tag)) +
  geom_line(lwd = 1)+
  ggtitle("How often was the 'R' and 'python' tag searched for?") +
  xlab("Year") +
  ylab("Percentage of all searches by year") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position = c(.20, .95),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.box.margin = margin(4, 4, 4, 4),
        legend.title.align=0.5)+
  labs(color = 'Tags')
```

#### **Let's find out how often R/Python libraries were searched**

```{r}
# A vector of selected tags 
selected_tags <- c('dplyr', 'ggplot2', "tidyverse", 'matplotlib', 'pandas', 'numpy')

# Filter for those tags
selected_tags_over_time <- tag_by_year_percentage %>%
  filter(tag %in% selected_tags)

head(selected_tags_over_time, 50)
```

###### Visualize it

```{r}
# Plot tags over time on a line plot using color to represent tag
ggplot(selected_tags_over_time, aes(year, fraction, color = tag)) +
  geom_line(size = 1, alpha = 0.5) +
  ggtitle("How often were R libraries searched?") +
  xlab("Year") +
  ylab("Percentage of searches") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position = c(.25, 1),
        legend.justification = c("right", "top"),
        legend.box.just = "right",
        legend.title.align=0.5)+
  labs(color = 'Tags')
```

#### **Excel and more...**

Let's see how often other data analysis programs such as 'spss', 'excel', 'sas', 'tableau', 'sql' were searched.

```{r}
analytisc_tags <- c("r", "python", "spss", 'excel', 'sas', 'tableau', 'sql')

# Filter for those tags
by_tag_subset <- tag_by_year_percentage %>%
  filter(tag %in% analytisc_tags)

by_tag_subset
```

###### Visualize it
```{r}
# Plot tags over time on a line plot using color to represent tag
ggplot(by_tag_subset, aes(year, fraction, color = tag)) +
  geom_line(size = 1, alpha = 0.5) +
  ggtitle("How often were data analysis programmes searched?") +
  xlab("Year") +
  ylab("Percentage of searches") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.title.align=0.5,
        legend.title = element_text(face = "bold"))+
  labs(color = 'Tags')
```

#### **What we were looking for most**

Finally, let's find out what were the most searched tags on stack overflow between 2008 and 2018

```{r}
sorted_tags <- tag_by_year %>%
  group_by(tag) %>%
  summarize(tag_total = sum(number)) %>%
  arrange(desc(tag_total))

# Print the new table
print(sorted_tags)
```

On the search podium were 'javascript', 'java' and 'c#'. Let's see how the race has looked over these 10 years

```{r}
# Get the six largest tags
highest_tags <- head(sorted_tags$tag)

# Filter for the six largest tags
by_tag_subset <- tag_by_year_percentage %>%
  filter(tag %in% highest_tags)

# Plot tags over time on a line plot using color to represent tag

ggplot(by_tag_subset, aes(year, fraction, color = tag)) +
  geom_line(size = 1, alpha = 0.5) +
  ggtitle("Most searched programming languages between 2008 - 2018") +
  xlab("Year") +
  ylab("Number of searches") +
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.title.align=0.5,
        legend.title = element_text(face = "bold"))+
  labs(color = 'Tags')


```


